---
title: "MATH40005 Coursework Spring 2021"
subtitle: "Xiao Xuan Tan, CID: 01938572"
output: pdf_document
---

<!--
This is a comment.
To create the pdf in RStudio, click on the "Knit" icon.
-->

# Introduction 

Testing if there is a statistically significant difference between the average 
heights of the people in countries X and Y.




# Question 1

Read in the data.

```{r}
df1<-read.table("~/Desktop/MATH40005_coursework_2021_questions/x_data.txt", 
                sep=",",header=T)
df2<-read.table("~/Desktop/MATH40005_coursework_2021_questions/y_data.txt", 
                sep=",",header=T)
x<-df1$x
y<-df2$y
```

# Question 2

This box plot compares the median, quartiles, maximum and minimum of the 
heights in country X and Y.

<!--
This block is a comment, below is a code chunk.
Setting the out.width option specifies how large to make any plots.
You can increase/decrease the size of the plot by changing the value 0.7
-->


```{r, out.width="0.675\\linewidth"}
boxplot(x,y,horizontal=TRUE,names=c("Country X","Country Y"),xlab="Heights",
        main="Comparison of heights in country X and country Y")
```

<!--
Use the command below to start a new page
-->
\clearpage




# Question 3

Suppose n people are randomly selected in country X and their heights are the 
random variables $X_1, X_2, \dots, X_n$, which are assumed to be independent and 
identically distributed following a normal distribution with unknown mean $\theta_1$ 
and unknown variance $\sigma^2_1$.
The observations are $x_1, x_2, \dots, x_{n}$.

Suppose m people are randomly selected in country Y and their heights are the 
random variables $Y_1, Y_2, \dots, Y_m$, which are assumed to be independent and 
identically distributed following a normal distribution with unknown mean $\theta_2$ 
and unknown variance $\sigma^2_2$.
The observations are $y_1, y_2, \dots, y_{m}$.

Assume that $\sigma^2_1$=$\sigma^2_2$.

The null hypothesis is:  

  - $H_0$: $\theta_1=\theta_2$
 
The alternative hypothesis is:  

  - $H_1$: $\theta_1\neq\theta_2$
 
I plan to use the student's two sample test to test the hypothesis.
The significance threshold is $\alpha=0.05$.



# Question 4

Assumption 1: 
Most of the points in both of the Q-Q plots lie along the line y = x, and so X 
and Y can each be assumed to be normally distributed.

```{r, out.width="0.7\\linewidth"}
layout(matrix(c(1,2),nrow=1,ncol=2,byrow=FALSE))
z<-(x-mean(x))/sd(x)
qqnorm(z,main="Normal Q-Q Plot of X")
abline(0,1,col="red")
w<-(y-mean(y))/sd(y)
qqnorm(w,main="Normal Q-Q Plot of Y")
abline(0,1,col="red")
```

Assumption 2: 
The ratio of the standard deviation of x and y is approximately 1, thus the variances 
of X and Y can be assumed to be equal.
```{r}
cat("The standard deviation of x is: ", sd(x), "\n", sep="")
cat("The standard deviation of y is: ", sd(y), "\n", sep="")
cat("The ratio of the standard deviations of x and y is: ",sd(y)/sd(x), "\n", sep="")
```

# Question 5
The t-statistic is 
$$T=\frac{(\bar{X}-\bar{Y})-(\theta_1-\theta_2)}{s_{p} \sqrt{\frac{1}{n}+\frac{1}{m}}}$$
where 
$$s_{p}^2 = \frac {(n - 1)s_{1}^2 + (m - 1)s_{2}^2}{n + m - 2}$$ 
is the pooled sample variance.

```{r}
n<-length(x)
m<-length(y)
x_bar<-mean(x)
y_bar<-mean(y)
x_var<-var(x)
y_var<-var(y)
pooled_variance<-((n-1)*x_var+(m-1)*y_var)/(n+m-2)
t<-(x_bar-y_bar)/(sqrt(pooled_variance)*sqrt(1/n+1/m))
cat("The test statistic t=", t, "\n", sep="")

alpha<-0.05
nu<-n+m-2
c<-qt(1-alpha/2,nu)
cat("The critical threshold is |t|>",c,"\n",sep="")
```

\clearpage
# Question 6

```{r}
cat("Since the realised statistic |t|=",abs(t),">",c,", the null hypothesis is rejected.", 
    "\n",sep="")
```
The data supports the case that there is a statistically significant difference 
between the average heights of people in countries X and Y.

